Automatic Extraction of JavaScript Formal Semantics from ECMAScript Specifications

### 2019-08-06 icse/compiler.tex
Renamed converter -> compiler

\section{Algorithm Converter}

There are several efforts to extract formal ECMAScript semantics for static analysis
or verifications of JavaScript programs~\cite{???}. However, existing researches
manually extract the semantics and only target specific version of
EMCAScript specifications, mostly ECMAScript 5.1.
It is the critical problem to treat real-world JavaScript programs
because ECMAScript specifications are annually updated after 2015.

To alleviate such problems, we propose to automatically extract semantics from
ECMAScript specifications. Their semantics are provided as \textit{abstract algorithms}
written in structural natural langauges. Our main idea is to convert such algorithms
into our small core languages. The algorithm converter consits of multiple rules
for each layer. Moreover, we provide \textit{rule generation assistant}.
It diagnoses root causes of failed conversions and suggests
that which layer of rules should be extended for them.

@@

Abstract algorithms in ECMAScript specifications are written in structural
natural languages. For example, Figure~\ref{fig:to-primitive} describes
the \( \code{ToPrimitive} \) abstract algorithm. It takes two parameters
\( \code{input} \) and \( \code{PreferredType} \) and evaluates its body
with given arguments. Its body consits of multiple steps and each step could
have multiple sub-steps as well. While it is not easy to infer exact meaning
of natural languages, fortunately, there is additional HTML tags are attached
into several tokens in abstract algorithms. For example, all the paramters
and newly defined local identifiers have \( \code{<var>} \) tags.
We utilize such information to construct more precise semantics.
The following table describes the list of HTML tags we utilize
and their meaning:

\[
  \begin{array}{c|l}

    \text{tags} & \text{descriptions}\\\hline\hline
    \code{<code>} & \text{ECMAScript codes}\\\hline
    \code{<emu-const>} & \text{constant values}\\\hline
    \code{<emu-grammar>} & \text{productions}\\\hline
    \code{<emu-nt>} & \text{non-terminal syntax}\\\hline
    \code{<emu-val>} & \text{values}\\\hline
    \code{<ol>} & \text{ordered sub-steps}\\\hline
    \code{<sup>} & \text{superscripts}\\\hline
    \code{<ul>} & \text{unordered sub-steps}\\\hline
    \code{<var>} & \text{variables}\\\hline
  \end{array}
\]

@@

We convert abstract algorithms

@@

\begin{figure}
  \centering
  \includegraphics[height=18em]{img/to_primitive.png}
  \begin{lstlisting}[style=myCorestyle]
"ToPrimitive" (input, PreferredType) => {
  if (= (Type input) "Object") {
    if (= PreferredType absent) let hint = "default"
    else if (= PreferredType "String") let hint = "string"
    else let hint = "number"
    let exoticToPrim = ? (GetMethod input SYMBOL_toPrimitive)
    if (! (= exoticToPrim undefined)) {
      let result = ? (Call exoticToPrim input (new [hint]))
      if (! (= (Type result) "Object")) return result
      return (Throw INTRINSIC_TypeErrorPrototype)
    }
    if (= hint "default") hint = "number"
    return ? (OrdinaryToPrimitive input hint)
  }
  return input
}
  \end{lstlisting}
  \caption{\( \code{ToPrimitive} \) abstract algorithms}
  \label{fig:to-primitive}
\end{figure}

>>>>

\section{Algorithm Compiler}

In this section, we propose a compiler from abstract algorithms
into our core language for ECMAScript specifications.
We first convert a given abstract algorithm into a list of
tokens based on HTML tags. Then, we

use \textit{compiler assistant}
to generate compiler rules.

o
In this section, we propose our core languages for ECMAScript
and explain how to convert abstract algorithms into the core programs.
We first explain the pre-processing to extract abstract algorithms
as lists of tagged tokens from ECMAScript specifications written in HTML.
Then, we introduce main features of the core languages and

@@

Abstract algorithms in ECMAScript specifications are written in structural
natural languages. For example, Figure~\ref{fig:to-primitive} describes
the \( \code{ToPrimitive} \) abstract algorithm. It takes two parameters
\( \code{input} \) and \( \code{PreferredType} \). The body consits of
multiple steps and each step might have multiple sub-steps as well.
While it is not easy to infer exact meaning of natural languages,
fortunately, there is additional HTML tags are attached
into several tokens in abstract algorithms. For example, all the paramters
and local identifiers have \( \code{<var>} \) tags. We utilize such information
to construct more precise semantics. The following table describes
the list of HTML tags we utilize and their meaning:

\[
  \begin{array}{c|l}
    \text{tags} & \text{descriptions}\\\hline\hline
    \code{<code>} & \text{ECMAScript codes}\\\hline
    \code{<emu-const>} & \text{constant values}\\\hline
    \code{<emu-grammar>} & \text{productions}\\\hline
    \code{<emu-nt>} & \text{non-terminal syntax}\\\hline
    \code{<emu-val>} & \text{values}\\\hline
    \code{<ol>} & \text{ordered sub-steps}\\\hline
    \code{<sup>} & \text{superscripts}\\\hline
    \code{<ul>} & \text{unordered sub-steps}\\\hline
    \code{<var>} & \text{variables}\\\hline
  \end{array}
\]

@@

We use multi-layer rules to convert abstract algorithms.
Each layer of rules denotes same level of expressiveness.
We define six different layers; statements, expressions,
conditions, types, values, references.

\paragraph{Statements}

@@

\begin{figure}
  \centering
  \includegraphics[height=18em]{img/to_primitive.png}
  \begin{lstlisting}[style=myCorestyle]
"ToPrimitive" (input, PreferredType) => {
  if (= (Type input) "Object") {
    if (= PreferredType absent) let hint = "default"
    else if (= PreferredType "String") let hint = "string"
    else let hint = "number"
    let exoticToPrim = ? (GetMethod input SYMBOL_toPrimitive)
    if (! (= exoticToPrim undefined)) {
      let result = ? (Call exoticToPrim input (new [hint]))
      if (! (= (Type result) "Object")) return result
      return (Throw INTRINSIC_TypeErrorPrototype)
    }
    if (= hint "default") hint = "number"
    return ? (OrdinaryToPrimitive input hint)
  }
  return input
}
  \end{lstlisting}
  \caption{\( \code{ToPrimitive} \) abstract algorithms
  and its generated core program.}
  \label{fig:to-primitive}
\end{figure}
